### 2022-01-18 documentos/math.tex
Fix the fixes

\section{Mathematical considerations} 
% Intro {{{
This section is devoted to derive the conditions a Pauli map needs to satisfy 
the complete positivity condition i.e. that all the eigenvalues of the Choi matrix
associated to the channel are non-negative.  To do so, we calculate
and diagonalize the Choi matrix associated with general Pauli maps, first for a
single qubit and then for $N$ qubits. We then restrict to PCEs and, for quantum
channels, associate a vector space to the set of coefficients
$\{\tau_\valpha\}$ in \eref{eq:PCE_definition}, from which we are able to
derive several important conclusions. 
% }}}
\subsection{Diagonalization of Choi matrix} % {{{

\flnote{
We now construct the Choi matrix associated to the single qubit map
$\mathcal{E}$. As described above, $\mathcal{E}$ is a linear map from
$\mcB(\hilbert)$ to itself. We shall denote elements of
$\mcB(\hilbert)$ by the notation $\kket{\cdot}$. Thus, for instance,
$\kket{\sigma_\alpha}$ represents the Pauli matrix $\sigma_\alpha$ understood
as a vector belonging to $\mcB(\hilbert)$, for the present case, in which
$\hilbert={\mathbb{C}^2}$. Since the scalar product on $\mcB(\hilbert)$ is
given by  $\bbrakket{A_1}{A_2}=\tr A_1^\star A_2$,
} 
% We now study the matrix form of the single qubit map $\mathcal{E}$ in
% order to construct the associated Choi matrix. 
% To do so, we shall denote operators in the superoperator space with the symbol
% $\kket{\cdot}$. For example, $\kket{\sigma_{\alpha}}$ correspond to the
% vectorized 
elements of the Pauli basis and satisfy the relation 
$\bbrakket{{\sigma_{\alpha}} }{{\sigma_{\alpha'} }} = \tr
\left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2
\delta_{\alpha\alpha'}$.
In this language, the state reads
$\kket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \kket{\sigma_{\alpha}}$
and the matrix form of the map $\mathcal{E}$ is
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha}
\ddyada{\sigma_{\alpha}}.
\end{equation}
After some steps, detailed from \eref{eq:inicio:canal:pauli} to
\eref{Choi_Pauli_1q}, it is possible to show that the associated Choi matrix
reads
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha}
\otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
Notice that $\ddyada{\sigma_{\alpha}}$ 
and $\sigma_{\alpha} \otimes \sigma_{\alpha}^*$ are different operators.
Indeed, the former acts as a linear map upon the vector space $\mcB(\hilbert)$,
whereas the latter acts on the tensor product $\hilbert\otimes\hilbert$. Of
course, there is a basis dependent identification between these two
spaces, which is used 
in the construction of the Choi matrix~\cite{}. 
% \cpnote{Aca poner algo de que no es trivial y que se detalla en el apendice.}
% 
% Moreover, observe that $\mathcal{D}$ has the same form as the matrix
% representation of a random unitary channel, with Kraus operators given by
% $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}.
Surprisingly, one can in fact show that $\mathcal{D}$ is diagonal in the Pauli
basis (see Appendix \ref{sec:appendix:diagonalization}  for details).
% In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see
% Appendix for details),
%
% \begin{equation}
%  \mathcal{D} \kket{\sigma_{\alpha}}= \lambda_{\alpha} \kket{\sigma_{\alpha}},
% \end{equation}
% 
The eigenvalues are
\begin{equation}
 \lambda_{\alpha} = \frac12 \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
% or more compactly $\vec{\lambda} = \sa \vec{\tau}$, 
where 
\begin{equation}
\sa=\left(
\begin{array}{cccc}
1 &1 &1   &1   \\
 1 & 1  & -1&-1   \\
1  & -1  & 1 & -1\\
1&-1&-1&1  
\end{array}
\right).
% =\frac12\sa^{-1}.
\label{eq:1}
\end{equation}
We wish to add that one can replace $a$ with $H \otimes H$, with $H$ the
Hadamard matrix, and still diagonalize the same Choi matrix. This is due to the
fact that $a$ corresponds to a permutation of rows of  $H \otimes H$. However,
we chose \janote{to keep} $a$ \janote{as in \eqref{eq:1},} as some later considerations [see  \eref{eq:sigma_property}]
cannot be written easily in terms of $H \otimes H$.

>>>>

\section{Mathematical considerations} 
% Intro {{{
This section is devoted to derive the conditions a Pauli map needs to satisfy 
the complete positivity condition i.e. that all the eigenvalues of the Choi matrix
associated to the channel are non-negative.  To do so, we calculate
and diagonalize the Choi matrix associated with general Pauli maps, first for a
single qubit and then for $N$ qubits. We then restrict to PCEs and, for quantum
channels, associate a vector subspace to the set of coefficients
$\{\tau_\valpha\}$ in \eref{eq:PCE_definition}. This allows us  to
derive several important conclusions. 
% }}}
\subsection{Diagonalization of Choi matrix} % {{{

We now construct the Choi matrix associated to the single qubit map
$\mathcal{E}$. As described above, $\mathcal{E}$ is a linear map from
$\mcB(\hilbert)$ to itself. We shall denote elements of
$\mcB(\hilbert)$ by the notation $\kket{\cdot}$. Thus, for instance,
$\kket{\sigma_\alpha}$ represents the Pauli matrix $\sigma_\alpha$ understood
as a vector belonging to $\mcB(\hilbert)$, for the present case, in which
$\hilbert={\mathbb{C}^2}$. Since the scalar product on $\mcB(\hilbert)$ is
given by  $\bbrakket{A_1}{A_2}=\tr A_1^\star A_2$,
% We now study the matrix form of the single qubit map $\mathcal{E}$ in
% order to construct the associated Choi matrix. 
% To do so, we shall denote operators in the superoperator space with the symbol
% $\kket{\cdot}$. For example, $\kket{\sigma_{\alpha}}$ correspond to the
% vectorized 
elements of the Pauli basis and satisfy the relation 
$\bbrakket{{\sigma_{\alpha}} }{{\sigma_{\alpha'} }} = \tr
\left(\sigma_{\alpha}^{\dagger} \sigma_{\alpha'}\right) = 2
\delta_{\alpha\alpha'}$.
In this language, the state reads
$\kket{\rho}=2^{-1}\sum_{\alpha=0}^{3} r_{\alpha} \kket{\sigma_{\alpha}}$
and the matrix form of the map $\mathcal{E}$ is
% 
\begin{equation}
\hat{\mathcal{E}} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha}
\ddyada{\sigma_{\alpha}}.
\end{equation}
After some steps, detailed from \eref{eq:inicio:canal:pauli} to
\eref{Choi_Pauli_1q}, it is possible to show that the associated Choi matrix
reads
\begin{equation}
 \mathcal{D} = \frac{1}{2}\sum_{\alpha=0}^{3} \tau_{\alpha} \sigma_{\alpha}
\otimes \sigma_{\alpha}^*.
%  \label{Choi_1}
\end{equation}
Notice that $\ddyada{\sigma_{\alpha}}$ 
and $\sigma_{\alpha} \otimes \sigma_{\alpha}^*$ are different operators.
Indeed, the former acts as a linear map upon the vector space $\mcB(\hilbert)$,
whereas the latter acts on the tensor product $\hilbert\otimes\hilbert$. Of
course, there is a basis dependent identification between these two
spaces, which is used 
in the construction of the Choi matrix~\cite{}. 
% \cpnote{Aca poner algo de que no es trivial y que se detalla en el apendice.}
% 
% Moreover, observe that $\mathcal{D}$ has the same form as the matrix
% representation of a random unitary channel, with Kraus operators given by
% $K_{\alpha}=\sqrt{\tau_{\alpha}/2}\sigma_{\alpha}$ \cite{CHRUSCINSKI20131425}.
Surprisingly, one can in fact show that $\mathcal{D}$ is diagonal in the Pauli
basis (see Appendix \ref{sec:appendix:diagonalization}  for details).
% In this way the Choi matrix $\mathcal{D}$ is diagonal in the Pauli basis (see
% Appendix for details),
%
% \begin{equation}
%  \mathcal{D} \kket{\sigma_{\alpha}}= \lambda_{\alpha} \kket{\sigma_{\alpha}},
% \end{equation}
% 
The eigenvalues are
\begin{equation}
 \lambda_{\alpha} = \frac12 \sum_{\beta=0}^3\sa_{\alpha\beta}\tau_{\beta},
\end{equation}
% or more compactly $\vec{\lambda} = \sa \vec{\tau}$, 
where 
\begin{equation}
\sa=\left(
\begin{array}{cccc}
1 &1 &1   &1   \\
 1 & 1  & -1&-1   \\
1  & -1  & 1 & -1\\
1&-1&-1&1  
\end{array}
\right).
% =\frac12\sa^{-1}.
\label{eq:1}
\end{equation}
We wish to add that one can replace $a$ with $H \otimes H$, with $H$ the
Hadamard matrix, and still diagonalize the same Choi matrix. This is due to the
fact that $a$ corresponds to a permutation of rows of  $H \otimes H$. However,
we chose \janote{to keep} $a$ \janote{as in \eqref{eq:1},} as some later considerations [see  \eref{eq:sigma_property}]
cannot be written easily in terms of $H \otimes H$.

### 2022-01-18 documentos/pce.tex
Fix the fixes

We shall start exploring a single qubit scenario, where the most general 
density matrix can be written as
\begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
\rho=\frac12 {\sum_{\alpha=0}^3 r_\alpha \sigma_\alpha},
\end{equation}
with $\sigma_0= \one$, and $\sigma_{1,2,3}$ the usual Pauli matrices. 
Normalization requires that $r_0=1$ and the remaining $r_{1,2,3}$ 
form the Bloch vector. Consider the map that projects
each component 
in the following way,
\begin{equation}
r_\alpha \mapsto \tau_\alpha r_\alpha
\end{equation}
where $\tau_\alpha$ is either $1$ or $0$. Not every
operation like this is a quantum map; for example collapsing the entire Bloch
ball to the $xy$ plane (setting $\tau_3=0$ and the rest to 1) is not a quantum
channel. A direct evaluation of the complete positivity
conditions~\cite{zimansbook,davalosdivisibility}, 
\begin{align}
1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
1+\tau_\alpha-\tau_\beta-\tau_\gamma &\geq 0 \ \ \forall \alpha\neq \beta \neq \gamma,
\end{align}
where trace-preserving is already imposed ($\tau_0=1$), show that five out of
the eight 1-qubit PCE maps are quantum channels.
% 
% show that only five
% operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
These operations are the identity map, the completely depolarizing channel
($\rho \mapsto \one/2$); and completely bit, phase, and the bit-phase flip
channels~\cite{chuangbook}, 
% These operations are by definition diagonal in the
% Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat
% \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$
and can be pictured using one column tables showing the positions of $0$s and
$1$s, see~\fref{fig:one:qubit:examples}.

@@

In order to present and work the $N$ qubit case, it is useful 
to introduce 
the so-called \textit{Pauli strings}, defined as 
\begin{equation}
\sigma_{\vec \alpha}=\sigma_{\alpha_1}\otimes \sigma_{\alpha_2} \otimes \dots \otimes \sigma_{\alpha_N},
\label{eq:pauli_strings}
\end{equation}
where $\vec \alpha$ denotes a vector index $\qty(\alpha_1,\ldots,\alpha_N)$
and $\alpha_i=0,1,2,3$. These hermitian operators form an orthogonal 
basis in the space of operators acting on $N$ qubits. In fact, 
$\tr \sigma_{\vec \alpha} \sigma_{\vec \alpha'}=2^N\delta_{\valpha \valpha'}$
and 
$\tr \sigma_{\vec \alpha}= 2^{N}\delta_{\vec\alpha \vec 0}$.
% 
% Thus, given $N$ and the fact that $\tr
% \sigma_{\vec \alpha}=\delta_{\vec\alpha,\vec 0} 2^{N}$, it is clear that the
% basis $\left\{ 2^{N/2} \sigma_{\vec \alpha} \right\}$ is orthonormal in
% $\mcB(\hilbert_{2^N})$. The case with $N=1$ corresponds to the well-known
% \textit{Pauli basis},
% $1/\sqrt{2} \left\{ \one, \sigma_x,\sigma_y,\sigma_z \right\}$.

Similarly to the one qubit case, the density matrix $\rho$ of a system of $N$
qubits can be written using Pauli strings in the following way, 
\begin{equation}\label{eq:N_qubits_rho}
	\rho =\frac{1}{2^N} 
            \sum_\valpha r_\valpha \sigma_\valpha,
%             \quad r_{\vec0}=1,
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
%             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
	%r_{0,\ldots,0}=1,
\end{equation}
so $r_{\vec \alpha}=\langle \sigma_{\vec \alpha} \rangle=\tr \left(\rho
\sigma_{\vec \alpha} \right)$ is the coefficient corresponding to the expansion
of the density matrix in the normalized basis of Pauli strings. Again,
normalization of the state requires that $r_{\vec 0}=1$.
We shall refer to $\paulicomponents$ as the ``Pauli components'' of the density
matrix of a system of qubits.

@@

A Pauli component erasing (PCE) map is a map
%  Pauli diagonal operation
that either preserves or completely erases the Pauli components of a
density matrix. That is, 
% it transforms its Pauli components as 
\begin{equation} \label{eq:PCE_definition}
% \begin{split}
% 	\paulicomponents \mapsto \ \taus \paulicomponents,\hspace{.5cm}\\
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
	\paulicomponents \mapsto \taus \paulicomponents,\quad 
	\taus = 0,1.
% \end{split}
\end{equation}
For the operation to be trace preserving, it is required that $\tau_{\vec
0}=1$. 
It is worth noticing that, as for the single qubit case, not all PCE maps are
quantum operations. 
On the other hand, constructing and
evaluating the conditions for complete positivity is involved and is the main
problem addressed in this paper. 
If we refer to the map  $\paulicomponents \mapsto \taus \paulicomponents$, 
with arbitrary values of 
% but allow the 
$\taus$ (only restricted by complete positivity) as Pauli diagonal operations.

>>>>

We shall start exploring a single qubit scenario, where the most general 
density matrix can be written as
\begin{equation}
% \rho=\frac{\one +\vec r \cdot \vec \sigma}{2}.
\rho=\frac12 {\sum_{\alpha=0}^3 r_\alpha \sigma_\alpha},
\end{equation}
with $\sigma_0= \one$, and $\sigma_{1,2,3}$ the usual Pauli matrices. 
Normalization requires that $r_0=1$ and the remaining $r_{1,2,3}$ 
form the Bloch vector. Consider the map that projects
each component 
in the following way,
\begin{equation}
r_\alpha \mapsto \tau_\alpha r_\alpha
\end{equation}
where $\tau_\alpha$ is either $1$ or $0$ (trace preserving requires
that $\tau_0=1$). Not every
such operation  is a quantum map; for example collapsing the entire Bloch
ball to the $xy$ plane (setting $\tau_{1,2}=1$ and  $\tau_3=0$) is not a quantum
channel. A direct evaluation of the complete positivity
conditions~\cite{zimansbook,davalosdivisibility}, 
\begin{align}
1+\tau_1+\tau_2+\tau_3 &\geq 0, \nonumber \\
1+\tau_\alpha-\tau_\beta-\tau_\gamma &\geq 0 \ \ \forall \alpha\neq \beta \neq \gamma,
\end{align}
where trace-preserving is already imposed, show that five out of
the eight 1-qubit PCE maps are quantum channels.
% 
% show that only five
% operations are \pce:
% \janote{Derivado de mi comentario sobre el uso de ``PCE maps''
% también propongo que en vez de ``show that only five operations are \pce{}''
% se cambie por ``show that five out of the eight 1-qubit PCE maps are quantum
% channels''} 
These operations are the identity map, the completely depolarizing channel
($\rho \mapsto \one/2$); and completely bit, phase, and the bit-phase flip
channels~\cite{chuangbook}, 
% These operations are by definition diagonal in the
% Pauli basis~\cite{Ziman2005,davalosdivisibility}, \ie{} $\hat
% \mcE=\diag{\left(1,\tau_x, \tau_y, \tau_z \right)}$
and can be pictured using one column tables showing the positions of $0$s and
$1$s, see~\fref{fig:one:qubit:examples}.

@@

In order to present and develop the $N$ qubit case, it is useful 
to introduce 
the so-called \textit{Pauli strings}, defined as 
\begin{equation}
\sigma_{\vec \alpha}=\sigma_{\alpha_1}\otimes \sigma_{\alpha_2} \otimes \dots \otimes \sigma_{\alpha_N},
\label{eq:pauli_strings}
\end{equation}
where $\vec \alpha$ denotes a vector index $\qty(\alpha_1,\ldots,\alpha_N)$
and $\alpha_i=0,1,2,3$. These hermitian operators form an orthogonal 
basis in the space of operators acting on $N$ qubits. In fact, 
$\tr \sigma_{\vec \alpha} \sigma_{\vec \alpha'}=2^N\delta_{\valpha \valpha'}$
and 
$\tr \sigma_{\vec \alpha}= 2^{N}\delta_{\vec\alpha \vec 0}$.
% 
% Thus, given $N$ and the fact that $\tr
% \sigma_{\vec \alpha}=\delta_{\vec\alpha,\vec 0} 2^{N}$, it is clear that the
% basis $\left\{ 2^{N/2} \sigma_{\vec \alpha} \right\}$ is orthonormal in
% $\mcB(\hilbert_{2^N})$. The case with $N=1$ corresponds to the well-known
% \textit{Pauli basis},
% $1/\sqrt{2} \left\{ \one, \sigma_x,\sigma_y,\sigma_z \right\}$.

Similarly to the one qubit case, the density matrix $\rho$ of a system of $N$
qubits can be written using Pauli strings in the following way, 
\begin{equation}\label{eq:N_qubits_rho}
	\rho =\frac{1}{2^N} 
            \sum_\valpha r_\valpha \sigma_\valpha,
%             \quad r_{\vec0}=1,
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
%             \sum_{\alpha_1,\ldots,\alpha_N=0}^3 \paulicomponents
% 	\sigma_{\alpha_1}\otimes \ldots \otimes \sigma_{\alpha_N},
	%r_{0,\ldots,0}=1,
\end{equation}
so $r_{\vec \alpha}=\langle \sigma_{\vec \alpha} \rangle=\tr \left(\rho
\sigma_{\vec \alpha} \right)$ is the coefficient corresponding to the expansion
of the density matrix in the normalized basis of Pauli strings. Again,
normalization of the state requires that $r_{\vec 0}=1$.
We shall refer to $\paulicomponents$ as the {\it Pauli components} of the density
matrix of a system of qubits.

@@

A Pauli component erasing (PCE) map is a map
%  Pauli diagonal operation
that either preserves or completely erases the Pauli components of a
density matrix. That is, 
% it transforms its Pauli components as 
\begin{equation} \label{eq:PCE_definition}
% \begin{split}
% 	\paulicomponents \mapsto \ \taus \paulicomponents,\hspace{.5cm}\\
% 	\taus = 0,1, \hspace{1.5cm}	\tau_{0,\ldots,0}=1.
	\paulicomponents \mapsto \taus \paulicomponents,\quad 
	\taus = 0,1.
% \end{split}
\end{equation}
For the operation to be trace preserving, it is required that $\tau_{\vec
0}=1$. 
It is worth noticing that, as for the single qubit case, not all PCE maps are
quantum operations. 
On the other hand, constructing and
evaluating the conditions for complete positivity is involved and is the main
problem addressed in this paper. 
We shall refer to the map  $\paulicomponents \mapsto \taus \paulicomponents$, 
with arbitrary values of 
% but allow the 
$\taus$ (only restricted by complete positivity) as {\it Pauli diagonal operations}.
